Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-01-18 paper/main.tex
adjust text of group (& individual) results

% hypo a
We hypothesize[d] that whole-brain analyses will/would reveal increased
hemodynamic activity spatially constrained \todo{i.e. not activity all over the
place} to the PPA as functionally defined by a visual localizer experiment using
pictures done with the same subjects.  \citep{sengupta2016extension}.
% hypo b
Focussing on the [``cleaner operationalization'' of the] audio-description, we
further hypothesize[d] that an exclusively auditory stimulus can/could
substitute a block-design experiment using pictures to localize a ``visual
area'' in individual subjects.\todo{needs short statement on why one would even want that}
% results: group
On a group average level,
% operationalization in naturalistic stimuli
results demonstrate that our approach based on annotations [of two ecologically
more valid stimuli] to operationalize spatial perception lead to increased
hemodynamic activity in [roughly] the same brain regions the block-designed
visual localizer.\todo{interpretation?}
% anterior vs. posterior
On both a group and individual subject level, a comparison of results from the
three datasets also suggests functional differences between the anterior and
posterior PPA.
% what does it mean
The current data add evidence to the hypothesis that the PPA might consist of
submodules that work together during visual perception of landscapes (cf.
\citep{baldassano2013differential}).
%
Lastly, our results suggest that the PPA's anterior part can be localized in
individual subjects by contrasting high-level semantic information occurring in
an auditory naturalistic stimulus.

@@

% study goal
The goal of this study was to investigate if cues about the spatial layout
embedded in two naturalistic stimuli correlate with increased hemodynamic
activity in the PPA that was define by using a visual localizer experiment
within the same subjects \citep{sengupta2016extension}.
% what we did We
We annotated a \unit[2]{h} movie and its temporally aligned audio-description to
create GLM contrasts similar to previous studies that used conceptualized
stimuli.
% group level
On a group average level, we compare results gained from both naturalistic
stimuli to results gained from the visual localizer.
% individual level
On the level single subjects, we also assess localization performance of a
``visual area'' using the auditory stimulus.\todo{``localization'' performance}
% NeuroVault
Unthreholded $z$-maps of all contrasts on a group level as well as individual
results of the primary $t$-contrasts for both naturalistic stimuli can be found
at \href{https://neurovault.org/collections/KADGMGVZ/}{NeuroVault}.\todo{still
private collection}

@@

% intro
First on a group average level, we tested if results of a whole-brain analysis
based on the audio-description's annotation would lead to increased hemodynamic
activity in the PPA.

\subsubsection{group}
% AV intro
As conceptual/cross-modal control of the group average results, we also analyzed
data from the audio-visual movie by contrasting different kinds of movie cuts
independent(ly) from the frames' visual content following the cut.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to a setting that was not
depicted before to a recurring setting (\texttt{vse\_new > vpe\_old}) yielded
three significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises, parts of the intracalcarine
and cuneal cortex, the lingual gyrus and retrosplenial cortex, the occipital and
temporal fusiform gyrus and the parahippocampal cortex (reported from posterior
to anterior) in both hemispheres.
% LOC
Two additional homologous clusters are located in the superior lateral occipital
cortex (left bigger than right).

% AD
The primary $t$-contrast for the audio-description (\texttt{geo, groom} >
non-spatial noun categories) yielded six significant clusters (see Table
\ref{tab:res-ao-group1}, Figure \ref{fig:group-slices}).
% overlap with ROI
Two homologous clusters are located in the anterior part of the PPA group
overlap that we gained from the visual localizer experiment.
% in detail
Specifically, the clusters of the current analysis are at the borders of the
posterior parahippocampal cortex, the occipital and temporal fusiform gyrus and
lingual gyrus.
% precuneus
Two additional clusters are located in the ventral precuneus extending into the
retrosplenial cortex.
% LOC
Finally, two homologous clusters are located in the superior lateral occipital
cortex.
% concluding statement
\todo[inline]{any concluding statement here or merely report the clusters?}

@@

\todo[inline]{other reasons than localization performance?}

\todo[inline]{make focus on AD clearer}

% why individual results in the first place.
We also investigated [localization performance] of both naturalistic paradigms
on the level of individual subjects.\todo{depends on ``naturalistic stimulus for
localization''}
% results in both subject and group space
Time series of all subjects were analyzed both in group space and subject space.
 space
For better comparison of results across participants, we here report results
($z$-maps; Z>3.4, p<.05 cluster-corrected) of the analyses that was performed in
group space.
% ref to figure
Figure \ref{fig:subs-thresh-ppa} depicts thresholded $z$-maps of the primary AD
and AV $t$-contrast, and the outline of the individual PPAs
\citep{sengupta2016extension} overlayed on the MNI152 template (transversal
slice z=-11 for all subjects was chosen as it depicts voxels of significant
clusters in almost all subjects).
% individual z-maps -> dataset
Results of the analyses that were performed in each subject's subject-space
(unthresholded $z$-maps and thresholded $z$-maps with Z>2.3, p<.05 cluster
corrected) can be found at
\href{https://neurovault.org/collections/KADGMGVZ/}{NeuroVault}.\todo{still
private collection} .
% report here in group

@@

% PPA ROI (Sengupta et al., 2016)
The visual localizer experiment \citep{sengupta2016extension} yielded bilateral
clusters in 12 of 14 subjects and a unilateral right clusters in two subjects
(sub-04, sub-20) that were defined as the PPA.\todo{this is probably the part
that Simon considered as/too ``en passant''}
% procedure in Sengupta
Individual PPAs were assessed by visually inspecting unthresholded $z$-maps of
three different contrasts, and setting a subjectively best fitting $z$-threshold
to define an area in the parahippocampal gyrus [and adjacent regions of the
fusiform gyrus and anterior lingual gyrus].
% AD
In the current analysis of the AD stimulus, we use the same AD (and AV) contrast
for all subjects. We find statistically significant bilateral clusters in nine
subjects, and one cluster in the left hemisphere of sub-09 that are within or
overlapping with the individual PPA.
% subj-04
In sub-04, two homologous clusters are apparent, whereas the dedicated localizer
(and AV stimulus) yielded only one cluster in the right hemisphere.
% AV
Results of the AV contrast yield homologous clusters in five subjects,
unilateral right clusters in six subjects (of which one subjects yielded a
unilateral cluster in the visual localizer), and a unilateral left cluster in
one subject.
% difference to dedicated localizer
We find homologous clusters in sub-20, whereas the dedicated visual localizer
yielded only one cluster in the right hemisphere.
% concluding statement
\todo[inline]{concluding statement that does not sound like
interpretation/discussion?}

>>>>

% hypo a
We hypothesize[d] that whole-brain analyses will/would reveal increased
hemodynamic activity spatially constrained \todo{i.e. not activity all over the
place} to the PPA as functionally defined by a visual localizer experiment using
pictures done with the same subjects \citep{sengupta2016extension}.

We hypothesize[d] that an exclusively auditory stimulus can/could localize
``visual area'' in individual subjects and thus possibly substitute a
block-design experiment using pictures.
% hypo b
Further, we hypothesize[d] that an exclusively auditory stimulus can/could
localize the PPA in individual subjects, and thus substitute a visual experiment
be which would be more practicable to assess brain functions in visually
impaired persons.\todo{phrasing}
% results: group
On a group average level,
% operationalization in naturalistic stimuli
results demonstrate that our approach based on annotations [of two ecologically
more valid stimuli] to operationalize spatial perception lead to increased
hemodynamic activity in [roughly] the same brain regions the block-designed
visual localizer.\todo{interpretation?}
% anterior vs. posterior
On both a group and individual subject level, a comparison of results from the
three datasets also suggests functional differences between the anterior and
posterior PPA.
% what does it mean
The current data add evidence to the hypothesis that the PPA might consist of
submodules that work together during visual perception of landscapes (cf.
\citep{baldassano2013differential}).
%
Lastly, our results suggest that the PPA's anterior part can be localized in
individual subjects by contrasting high-level semantic information occurring in
an auditory naturalistic stimulus.

@@

% study goal
The goal of this study was to investigate if cues about the spatial layout
embedded in two naturalistic stimuli correlate with increased hemodynamic
activity in the PPA that was defined using a visual localizer experiment
within the same subjects \citep{sengupta2016extension}.
% what we did We
We annotated a \unit[2]{h} movie and its temporally aligned audio-description to
create GLM contrasts similar to previous studies that used static pictures.
% group level
On a group average level, we compare results gained from both naturalistic
stimuli to results gained from the visual localizer.
% individual level
On the level single subjects, we also assess the performance of a naturalistic
auditory stimulus as a substitute for a visual paradigm to localize the
PPA.\todo{``localize''}
% NeuroVault
Unthreholded $z$-maps of all contrasts on a group level as well as individual
results of the primary $t$-contrasts for both naturalistic stimuli can be found
at \href{https://neurovault.org/collections/KADGMGVZ/}{NeuroVault}.\todo{still
private collection}

@@

% intro
First, we tested if results of a whole-brain analysis based stimulus annotations
would lead to increased hemodynamic activity in the PPA.

\subsubsection{group}
% AV intro
First, we analyzed data from the audio-visual movie offering visual but
ecologically more valid stimulation.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to a setting that was not
depicted before to a recurring setting (\texttt{vse\_new > vpe\_old}) yielded
three significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises, parts of the intracalcarine
and cuneal cortex, the lingual gyrus and retrosplenial cortex, the occipital and
temporal fusiform gyrus and the parahippocampal cortex (reported from posterior
to anterior) in both hemispheres.
% LOC
Two additional homologous clusters are located in the superior lateral occipital
cortex (left bigger than right).

% AD intro
Second, we analyzed data from the movie's audio-description offering exclusively
auditory stimulation.
% AD primary contrast
The primary $t$-contrast for the audio-description (\texttt{geo, groom} >
non-spatial noun categories) yielded six significant clusters (see Table
\ref{tab:res-ao-group1}, Figure \ref{fig:group-slices}).
% overlap with ROI
Two homologous clusters are located in the anterior part of the PPA group
overlap that we gained from the visual localizer experiment.
% in detail
Specifically, the clusters of the current analysis are at the borders of the
posterior parahippocampal cortex, the occipital and temporal fusiform gyrus and
lingual gyrus.
% precuneus
Two additional clusters are located in the ventral precuneus extending into the
retrosplenial cortex.
% LOC
Finally, two homologous clusters are located in the superior lateral occipital
cortex.
% concluding statement
\todo[inline]{any concluding statement here or merely report the clusters?}

@@

\todo[inline]{imo no need to mention it here but mention in discussion that
results are not driven by outliers when discussing ``localization'' performance}

@@

% intro
Third, we investigated result from both naturalistic paradigms on the level of
individual subjects.
%
Focussing on the AD stimulus, wee hypothesize[d] that an exclusively auditory
stimulus can/could localize ``visual area'' in individual subjects and thus
possibly substitute a paradigm using blocks of pictures.
% AD results in
Time series of all subjects were analyzed both in group space and subject space.
For better comparions of subjects, we here report results from the analyses
performed in group space ($z$-maps; Z>3.4, p<.05 cluster-corrected).
% ref to figure
Figure \ref{fig:subs-thresh-ppa} depicts thresholded $z$-maps of the primary AD
and AV $t$-contrast, and the outline of overlapped, individual PPAs
\citep{sengupta2016extension}.
% MNI space
The $z$-maps and PPA outline are overlayed on the co-registered MNI152 template.
% one slice to show them all
Transversal slice of z=-11 was chosen as it depicts [at least some] voxels of
significant clusters in almost all subjects.
% individual z-maps -> dataset
Results of the analyses that were performed in each subject's subject-space
(unthresholded $z$-maps and thresholded $z$-maps, Z>2.3, p<.05 cluster
corrected) can be found at
\href{https://neurovault.org/collections/KADGMGVZ/}{NeuroVault}.\todo{still
private collection} .
% report here in group

\todo[inline]{following is probably the part that Simon considered as en
passant}

\todo[inline]{if we here report procedure of Sengupta as detailed as
it is now, we probably need to explain the differences between strict, relaxed,
simple in more detail in the method section and not just reference to Sengupta}

% Sengupta et al., 2016
The visual localizer experiment \citep{sengupta2016extension} yielded bilateral
[parahippocampal] clusters in 12 of 14 subjects and a unilateral right clusters
in two subjects (sub-04, sub-20).
% Sengupta: 1 of 3
Individual PPAs were assessed by visually inspecting thresholded $z$-maps
(Z>1.64, p<.05 cluster corrected) of three different $t$-contrasts
(\textit{strict}, \textit{relaxed}, and \textit{simple} stimulus set, cf.
\citep{sengupta2016extension}).
% Sengupta: procedure a)
Starting with the most conservative contrast (strict set) and a threshold of t =
2.5, we looked for clusters with at least 20 voxels (using AFNI).
%
We titrated [?] the threshold in the range of [2, 3] until we found an isolated
cluster for the PPA.
% Sengupta: procedure b)
If a cluster was not found or not isolated, we used a contrast from the relaxed
set, or finally the simple set, and repeated the process until we found a
cluster that matched the expected anatomical location based on literature
\citep{epstein1998ppa}.

% current analyses: one contrast to localize them all
In the current analyses, we use same contrast of each naturalistic stimulus for
all subjects

for all subjects. We find statistically significant bilateral clusters in nine
subjects, and one cluster in the left hemisphere of sub-09 that are within or
overlapping with the individual PPA.
% subj-04
In sub-04, two homologous clusters are apparent, whereas the dedicated localizer
(and AV stimulus) yielded only one cluster in the right hemisphere.
% AV
Results of the AV contrast yield homologous clusters in five subjects,
unilateral right clusters in six subjects (of which one subjects yielded a
unilateral cluster in the visual localizer), and a unilateral left cluster in
one subject.
% difference to dedicated localizer
We find homologous clusters in sub-20, whereas the dedicated visual localizer
yielded only one cluster in the right hemisphere.
% concluding statement
